Lab 8: Binomiaal Models in R

### 2023-01-13 Lab 10 PoissonNBmodels/Lab10_PoissonNBmodels_AnswerKey.Rmd
Add files via upload

#Predicted Counts

@@

>>>>

# Predicted Counts

@@

# Exercise

To do on your own:

Below, try using ggplot to plot this information we just generated. Follow these guidelines:
- X axis: Average number of Democratic Cliques
- Y axis: Predicted count of Intl. Wars
- Include a geom_line() for Dem_Cliques
- Include a geom_ribbon() with CIs (80% confidence interval)
- Include a legend which specifies what the line is and what the CIs are, try putting that legend on the bottom

### 2023-01-13 Lab 11 Missing data/Lab11_MissingData.Rmd
Add files via upload

```{r, eval = FALSE}
#Load in packages
pacman::p_load(mice,
               VIM) #Visualization

@@

summary(mice_data_2)
```

So this is great, but then what do we do if we want to run a model? We have five data sets. One option is to pool the data this way:
```{r, eval = FALSE}
#pooling
model_mice <- with(mice_data,lm(Temp~ Ozone + Solar.R + Wind))
summary(pool(model_mice))
```
So mice is great. Let's also learn about Amelia II (my personal favorite, named after Amelia Earhart, because she is missing). Amelia is a little different because it assumes that your data is jointly distributed as multivariate normal. It uses the "expectation-maximization with bootstrapping" as an algorithm for imputation. The EM algorithm alternates between something called the expectation (E-step) and maximization (M-step) steps until it converges on values that were missing - it converges when the current and previous values are quite close (sounds a bit like Newton Raphson, no?). It brings in bootstraping by doing this process on multiple bootstrapped samples drawn from your original (and incomplete) data.

@@

#Load in packages
pacman::p_load(Amelia,
               lubridate,
               xtable)#For tables

#create running time variable, lets just assume we're in 2021
data_miss$time <- make_datetime(month = data_miss$Month,
                                day = data_miss$Day,
                                year = 2023)

@@

#Pulling into DF
pooled_df <- data.frame(summary(pool(models_amelia)))

#Displaying in table
xtable(pooled_df)

>>>>

```{r, eval = FALSE}
#Load in packages
pacman::p_load(mice,
               VIM,
               texreg) #Visualization and tables

@@

summary(mice_data_2)

So this is great, but then what do we do if we want to run a model? We have five data sets. One option is to pool the data this way:
```{r, eval = FALSE}
#pooling
model_mice <- with(mice_data,lm(Temp~ Ozone + Solar.R + Wind))
summary(pool(model_mice))

#Using texreg package to display results in a table
texreg(pool(model_mice))

```
So mice is great. Let's also learn about Amelia II (my personal favorite, named after Amelia Earhart, because she is missing). Amelia is a little different because it assumes that your data is jointly distributed as multivariate normal. It uses the "expectation-maximization with bootstrapping" as an algorithm for imputation. The EM algorithm alternates between something called the expectation (E-step) and maximization (M-step) steps until it converges on values that were missing - it converges when the current and previous values are quite close (sounds a bit like Newton Raphson, no?). It brings in bootstraping by doing this process on multiple bootstrapped samples drawn from your original (and incomplete) data.

@@

#Load in packages
pacman::p_load(Amelia,
               lubridate)

#create running time variable, lets just assume we're in 2023
data_miss$time <- make_datetime(month = data_miss$Month,
                                day = data_miss$Day,
                                year = 2023)

@@

#Putting into table
texreg(pool(models_amelia))

### 2023-01-13 Lab 11 Missing data/Lab11_MissingData_AnswerKey.Rmd
Add files via upload

```{r, eval = FALSE}
#Load in packages
pacman::p_load(mice,
               VIM) #Visualization

@@

#try this out with 25 iterations and 10 imputations
#select animputation method (try pmm) - what does it do differently?
mice_data_2 <- mice(data_miss, #call our data
                  m=10, #number of multiple imputations
                  maxit=25, #number of iterations
                  method = "pmm",
                  seed=500)

summary(mice_data_2)

@@

@@

#create running time variable, lets just assume we're in 2021
data_miss$time <- make_datetime(month = data_miss$Month,
                                day = data_miss$Day,
                                year = 2023)

@@

#let's run the model again, but this time using a for loop
models_amelia <- list() #create with as many elements that you know exist
#create a vector of type list of size five

@@

@@

**On your own:**

@@

>>>>

```{r, eval = FALSE}
#Load in packages
pacman::p_load(mice,
               VIM,
               texreg) #Visualization

@@

#try this out with 25 iterations and 10 imputations
#select an imputation method (try pmm) - what does it do differently?
mice_data_2 <- mice(data_miss, #call our data
                  m=10, #number of multiple imputations
                  maxit=25, #number of iterations
                  method = "pmm",
                  seed=500)

summary(mice_data_2)
```

@@

#Using texreg package to display
texreg(pool(model_mice))

@@

#create running time variable, lets just assume we're in 2023
data_miss$time <- make_datetime(month = data_miss$Month,
                                day = data_miss$Day,
                                year = 2023)

@@

#let's run the model again, but this time using a for loop
models_amelia <- vector("list", 5)  #vector for our 5 newly imputed data sets

@@

#Putting into table
texreg(pool(models_amelia))

@@

#Exercise **On your own:**

@@

#Putting into table
texreg(pool(models_africa))

### 2023-01-13 Lab 6  Diagnostics and simulation based inferences/Lab6_Simulation_Uncertainty.Rmd
Add files via upload

# Simulation
```{r, message = FALSE}
#clear global environ
rm(list=ls())

@@

>>>>

# Simulation
```{r, message = FALSE}
#clear global environment
rm(list=ls())

@@

And last but not least, you can use the performance package (and the see package) to generate really nicely formatted regression assumption plots!
```{r}
pacman::p_load(see,
               performance)

#using the performance package to generate plots (zoom out if you get an error about plot size)
check_model(m1)

# performance package also has functions to help check assumptions
check_normality(m1)
?check_normality

check_heteroskedasticity(m1)
?check_heteroskedasticity

### 2023-01-13 Lab 6  Diagnostics and simulation based inferences/Lab6_Simulation_Uncertainty_Answers.Rmd
Add files via upload

# To do on your own:

>>>>

And last but not least, you can use the performance package (and the see package) to generate really nicely formatted regression assumption plots!
```{r}
pacman::p_load(see,
               performance)

#using the performance package to generate plots (zoom out if you get an error about plot size)
check_model(m1)

# performance package also has functions to help check assumptions
check_normality(m1)
?check_normality

check_heteroskedasticity(m1)
?check_heteroskedasticity

```
# To do on your own:

### 2023-01-13 Lab 8  BinomialModels/Lab8_BinomialModels.Rmd
Add files via upload

# Assignment

Read in the data 'turnout.csv.' Fit a logit model that predicts voting based on race, education and income.

Now, fit a second model that predicts voting based on race, education, income, and age. Assume that you've hypothesized that age has a curvilinear relationship with voting turnout, such that lower ages and higher ages are less likely to vote. Include this transformed variable.

Compare the model fit of these two options using lrtest, AIC, and BIC. Which do you choose as the better fitting model?

Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.

Generate an average predicted effect over the observed values of income for the first model. Plot this too.

@@

# Average-Case Approach
#When we combine the average-case approach with simulation, we create a hypothetical design matrix, where we hold all continuous variables at their mean or another justifiable value, and then let the predictor of interest vary along some range.

@@

# Observed-Case Approach
#When taking the observed-case approach (see Hanmer and Kalkan (2013) for why this is the preferable approach) and using simulation to get confidence intervals, we have to do things slightly differently. Specifically, we will have to take an addition mean.

@@

#
colnames(______) <- c('pe', 'lo', 'hi', 'income')

>>>>

# Exercise

+ Read in the data 'turnout.csv.' Fit a logit model that predicts voting based on race, education and income.

+ Now, fit a second model that predicts voting based on race, education, income, and age. Assume that you've hypothesized that age has a curvilinear relationship with voting turnout, such that lower ages and higher ages are less likely to vote. Include this transformed variable.

+ Compare the model fit of these two options using lrtest, AIC, and BIC. Which do you choose as the better fitting model?

+ Average-Case Approach: Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.

+ Observed-Case Approach: Generate an average predicted effect over the observed values of income for the first model. Plot this too.

@@

#Which is better?

# Average-Case Approach

#When we combine the average-case approach with simulation, we create a hypothetical design matrix, where we hold all continuous variables at their mean or another justifiable value, and then let the predictor of interest vary along some range.

@@

# Observed-Case Approach

#When taking the observed-case approach (see Hanmer and Kalkan (2013) for why this is the preferable approach) and using simulation to get confidence intervals, we have to do things slightly differently. Specifically, we will have to take an addition mean.

@@

colnames(______) <- c('pe', 'lo', 'hi', 'income')

ggplot(______)

### 2023-01-13 Lab 8  BinomialModels/Lab8_BinomialModels_AnswerKey.Rmd
Add files via upload

# Assignment

Read in the data 'turnout.csv.' Fit a logit model that predicts voting based on race, education and income.

Now, fit a second model that predicts voting based on race, education, income, and age. Assume that you've hypothesized that age has a curvilinear relationship with voting turnout, such that lower ages and higher ages are less likely to vote. Include this transformed variable.

Compare the model fit of these two options using lrtest, AIC, and BIC. Which do you choose as the better fitting model?

Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.

Generate an average predicted effect over the observed values of income for the first model. Plot this.

>>>>

# Exercise

+ Read in the data 'turnout.csv.' Fit a logit model that predicts voting based on race, education and income.

+ Now, fit a second model that predicts voting based on race, education, income, and age. Assume that you've hypothesized that age has a curvilinear relationship with voting turnout, such that lower ages and higher ages are less likely to vote. Include this transformed variable.

+ Compare the model fit of these two options using lrtest, AIC, and BIC. Which do you choose as the better fitting model?

+ Average-Case Approach: Simulate and plot predicted probabilities over the range of age, holding all other variables at their mean, for the second model. Plot this.

+ Observed-Case Approach: Generate an average predicted effect over the observed values of income for the first model. Plot this too.
